---
title: "Project Work"
format: html
editor: visual
---

## Project Work

This notebook presents the project work developed as part of the course **"Text Analysis and Large Language Models for Innovation Management"**, held on **September 1/2/3/4** in Pisa. The course was carried out by **Vito Giordano**, **Filippo Chiarello**, and **Ivan Zupic**.

The objective of this notebook is to provide guidelines in the application of **Natural Language Processing (NLP)** techniques to explore innovation-related topics, as introduced and practiced during the course. It includes the different **assignments and exercises** covered throughout the **hands-on sessions**.

By applying the methods discussed in class to a real-world case study, this work aims to demonstrate how textual data can be used to extract insights relevant to **innovation management** and decision-making.

## Case Study: Circular Economy Papers

In recent years, the **circular economy** has gained significant attention as a sustainable model aimed at reducing waste and optimizing resource use through innovative practices. Companies and organizations across diverse industries are increasingly adopting circular economy principles to redesign products, processes, and business models. This surge of interest is reflected in the growing body of **research articles** focused on circular economy innovations, highlighting new strategies and technologies that enable sustainable value creation.

Effective **innovation management** is vital for organizations seeking to implement circular economy practices successfully. By strategically managing innovation, businesses can enhance their sustainability performance while gaining a competitive advantage in evolving markets. In this framework, analyzing the literature on **circular economy innovation practices** offers valuable insights into emerging trends, best practices, and challenges, helping stakeholders anticipate future developments and drive systemic change toward more sustainable production and consumption patterns.

The hands-on component of this project is organized into **five assignments**, each focusing on a specific aspect of text analysis applied to the topic of **circular economy papers**:

-   **Assignment 1**: *Text Pre-processing of* *Circular Economy Papers*

-   **Assignment 2**: *Topic Modeling on Circular Economy Papers*

-   **Assignment 3**: *Sentiment Analysis on Circular Economy Papers*

-   **Assignment 4**: *Extracting Innovation Practices in Circular Economy Papers*

-   **Assignment 5**: *Mining* *Insights from Textual Data*

## **Assignment 1**: *Text Pre-processing of* *Circular Economy Papers*

The dataset contains a total of \[insert number\] articles (including titles, abstracts, and keywords) collected from the Scopus database. The data were extracted by using the following query:

`( TITLE-ABS-KEY ( ( "circular economy" OR "ecodesign" OR "eco design" OR "industrial symbiosis" OR "waste prevention" OR "waste reduction" OR "secondary raw material" OR "resource recovery" OR "sustainability criteria" OR "bioeconomy" OR "bio economy" OR "biorefinery" OR "bio refinery" OR "renewable resource" OR "waste recycling" ) ) AND TITLE ( innovat* ) )`

The resulting dataset is stored in a **CSV file** and includes:

-   The **title**, **abstract**, and **keywords** of each scientific article;

-   **Bibliographic metadata**, such as authors, publication year, source title, and affiliations.

The main goal of this assignment is to **explore and analyze the textual content** of circular economy papers, focusing specifically on Part-of-Speech (PoS) tagging and word count.

To conduct this analysis, you will apply a selection of Natural Language Processing (NLP) techniques introduced in the previous course modules. The workflow includes: (1) Text Preprocessing (Tokenization, PoS tagging, Stopword removal, Lemmatization or Stemming); and (2) Basic text analysis through word frequency counts and PoS distribution.

This assignment serves as a preliminary exploration of the dataset, aiming to provide initial insights into the structure and linguistic patterns of the job postings content. It is intended as a first step before moving on to more advanced analyses in subsequent assignments.

```{r}
library(tidyverse)

# Import Scientific Articles
circular_economy_papers <- read_csv("data/Paper Circula economy and Innovation.csv")

# Have a glimpse in the dataset
glimpse(circular_economy_papers)

# Main Variables
circular_economy_papers %>% 
  select(EID, Title, Abstract, `Index Keywords`, `Author Keywords`, Year, Authors, `Source title`) %>% 
  head(5)
```

*!!! Now it's your turn:*

**Assignment 1** – Perform a word count to identify the most frequent terms, then apply PoS tagging to extract the most common bigrams and nouns, gaining an initial overview of the circular economy papers.

```{r}

```

## **Assignment 2**: *Topic Modeling on Circular Economy Papers*

Building on the insights gained from Assignment 1, the focus of this task shifts from basic text preprocessing to deeper analysis through **topic modeling**.

The goal of this assignment is to uncover the **main innovation themes** emerging in circular economy research. By applying topic modeling techniques to a dataset of circular economy-related academic papers, you will identify which **topics, methods, and application areas** are most frequently addressed by researchers, offering a data-driven view of how innovation practices in the circular economy are evolving.

Use the cleaned and preprocessed data from Assignment 1 to: **(1) Apply topic modeling** to extract key topics from the circular economy paper dataset; **(2) Identify and visualize** the most representative terms for each topic; and **(3) Interpret and label** each topic in terms of innovation themes, application domains, or research directions.

*!!! Now it’s your turn:*

**Assignment 2** – Based on the processed dataset from Assignment 1, apply topic modeling to uncover the main themes in circular economy research papers and interpret the results.

```{r}

```

## **Assignment 3**: *Sentiment Analysis on Circular Economy Papers*

After exploring linguistic patterns and key topics in circular economy research papers, this assignment focuses on analyzing the **emotional tone** of innovation narratives through **sentence-level sentiment analysis**.

The goal is to assess how positively or negatively circular economy innovations are framed in academic texts by measuring the **distribution of sentiment** across titles and abstracts. This allows us to understand the tone of scholarly discourse, detect areas of enthusiasm or concern, and identify whether certain innovations or practices are presented more optimistically or cautiously.

In this assignment, you will: **(1) Apply sentiment analysis** techniques to classify each sentence from the paper titles and abstracts as positive, negative, or neutral; **(2) Analyze the overall sentiment** distribution to evaluate how circular economy innovations are described; **(3) Identify and examine the most positively and negatively framed sentences** to understand the narratives behind the extremes in sentiment.

*!!! Now it’s your turn:*

**Assignment 3** – Apply sentiment analysis to the sentences from circular economy research papers to assess their polarity, and identify the most positive and most negative sentences.

```{r}

```

## **Assignment 4**: *Extracting Innovation Practices in Circular Economy Papers*

In this assignment, we shift from general text analysis to a more focused extraction task aimed at identifying **key innovation practices mentioned** in circular economy research papers.

The goal is to extract and analyze the main practices and approaches discussed in titles and abstracts by applying a **Named Entity Recognition (NER)** approach using a gazetteer specific to circular economy innovation practices. This will help us understand which methods, strategies, or interventions are central to innovation in this field, and highlight the patterns shaping the evolution of circular economy research.

To simplify the NER task, we will use a **gazetteer-based approach**, relying on a curated list of relevant innovation practices.

Follow these steps: (1) **Prepare a Gazetteer** – Create a list of innovation practices commonly associated with the circular economy (e.g., product-service systems, industrial symbiosis, ecodesign, reverse logistics); (2) **Apply Gazetteer-based NER** – Use the gazetteer to identify and extract mentions of these practices from the titles and abstracts of the circular economy articles; (3) **Analyze the Results** – Count how often each practice appears, identify the most frequently mentioned ones, and visualize the results with tables or charts.

*!!! Now it’s your turn:*

**Assignment 4** – Create a gazetteer of innovation practices in the circular economy, extract them from research articles using Named Entity Recognition, and analyze which ones occur most frequently.

```{r}

```

## **Assignment 5**: *Mining* *Insights from Textual Data*

After exploring the key themes and innovation practices emerging from circular economy research, we now aim to gain deeper insights by addressing **key research questions** about how innovation is framed and discussed in this field, based on the analyses previously performed.

The goal of this assignment is to **select one** of the following research questions and use the techniques learned—such as text preprocessing, topic modeling, sentiment analysis, and entity extraction—to provide a data-driven answer. You are only expected to **design a clear plan** outlining how you would approach the analysis, and only **try to implement it [if you have time]{.underline}**.

1.  Which innovation practices are most frequently mentioned in circular economy papers, and how have they evolved over time?

2.  What are the main sectors or domains where circular economy innovations are being applied, as reflected in the research literature?

Think about the tools and methods introduced during the course to: **(1) Analyze the article dataset** with respect to your chosen question; **(2) Extract relevant insights** supported by both quantitative and qualitative evidence; **(3) Summarize your findings** with clear interpretations and visualizations that highlight the main takeaways.

*!!! Now it’s your turn:*

**Assignment 5** – Choose [one]{.underline} research question related to smartphone user reviews, and **design a clear plan** outlining how you would approach the analysis.

```{r}

```
