---
title: "Project Work"
format: html
editor: visual
---

## Project Work

This notebook presents the project work developed as part of the course **"Text Analysis for Innovation Management"**, held on **July 3/4/5** in Pisa. The course was carried out by **Vito Giordano**, **Filippo Chiarello**, **Ivan Zupic** and **Irene Spada**.

The objective of this notebook is to provide guidelines in the application of **Natural Language Processing (NLP)** techniques to explore innovation-related topics, as introduced and practiced during the course. It includes the different **assignments and exercises** covered throughout the **hands-on sessions**.

By applying the methods discussed in class to a real-world case study, this work aims to demonstrate how textual data can be used to extract insights relevant to **innovation management** and decision-making.

## Case Study: Smartphone User Reviews

In the context of product design and development, understanding **customer needs and preferences** is a crucial factor in ensuring market success. During the development phase, companies must be able to gather and process detailed information about consumer desires and expectations in order to create products that effectively meet those needs. This is especially true in a highly competitive market such as that of **smartphones**, where companies must respond quickly to consumer preferences and constantly evolving technological trends.

An important source of data for analyzing customer needs comes from **user reviews on e-commerce platforms** like Amazon. These reviews, often rich in detail, provide direct insight into user experiences, highlighting both the strengths and weaknesses of products. Analyzing this content allows companies to identify the most valued and useful product features, as well as pinpoint potential weaknesses or areas for improvement.

The hands-on component of this project is organized into **five assignments**, each focusing on a specific aspect of text analysis applied to the topic of **user-reviews of smartphones on e-commerce platforms**:

-   **Assignment 1**: *Text Pre-processing of* *User-reviews on Smartphones*

-   **Assignment 2**: *Topic Modeling on User-reviews*

-   **Assignment 3**: *Sentiment Analysis on User-reviews*

-   **Assignment 4**: *Extracting Smarphone Components in User-reviews*

-   **Assignment 5**: *Mining* *Insights from Textual Data*

## Assignment 1: *Text Pre-processing of* *User-reviews on Smartphones*

The dataset was downloaded directly from Kaggle at the following link:

https://www.kaggle.com/datasets/grikomsn/amazon-cell-phones-reviews

The dataset contains descriptions of various smartphones along with consumer reviews. It includes only the top 10 smartphone brands: ASUS, Apple, Google, HUAWEI, Motorola, Nokia, OnePlus, Samsung, Sony, and Xiaomi.

The resulting dataset is stored in a **CSV file** and includes:

-   The **title**, **body**, and **rating** of each user-review;

-   **Metadata**, such as asin, name, date, and helpful votes.

The main goal of this assignment is to **explore and analyze the textual content** of user reviews on smartphones, focusing specifically on word count and Part-of-Speech (PoS) tagging.

To conduct this analysis, you will apply a selection of Natural Language Processing (NLP) techniques introduced in the previous course modules. The workflow includes: (1) Text Preprocessing (Tokenization, PoS tagging, Stopword removal, Lemmatization or Stemming); and (2) Basic text analysis through word frequency counts and PoS distribution.

This assignment serves as a preliminary exploration of the dataset, aiming to provide initial insights into the structure and linguistic patterns of the user-generated content. It is intended as a first step before moving on to more advanced analyses in subsequent assignments.

```{r}
library(tidyverse)

# Import Scientific Articles
smartphone_reviews <- read_csv("data/Reviews.csv")

# Have a glimpse in the dataset
glimpse(smartphone_reviews)

# Main Variables
smartphone_reviews %>% 
  select(asin, name, rating, date, title, body, helpfulVotes) %>% 
  head(7)
```

*!!! Now it's your turn:*

**Assignment 1** ˇV Perform a word count to identify the most frequent terms, then apply PoS tagging to extract the most common bigrams and nouns, gaining an initial overview of the smartphone user reviews.

```{r}

```

## **Assignment 2**: *Topic Modeling on User-reviews*

Building on the insights gained from **Assignment 1**, the focus of this task shifts from basic text preprocessing to deeper analysis through **topic modeling**.

The goal of this assignment is to uncover the **main themes and concerns** expressed by consumers by applying topic modeling techniques to the review texts. This will allow you to identify recurring topics, understand user priorities, and highlight which product features or issues are most frequently discussed.

Use the cleaned and preprocessed data from Assignment 1 to: (1) Apply topic modeling to detect dominant topics in the reviews; (2) Show the most representative words for each topic, and (3) Interpret and label each topic based on them.

*!!! Now it's your turn:*

**Assignment 2** ˇV Based on the processed reviews from Assignment 1, apply topic modeling to uncover the main themes in user feedback on smartphones and interpret the results.

```{r}

```

## **Assignment 3**: *Sentiment Analysis on User-reviews*

ndefinedAfter exploring linguistic patterns and key topics in user reviews, this assignment focuses on analyzing the **emotional tone** of the feedback through a **review-level** **sentiment analysis**.

The goal is to assess **how polarized the reviews are** by measuring the distribution of **positive, negative, and neutral sentiments**. This helps us understand the overall satisfaction of users, detect strong emotional responses, and identify which products or features tend to trigger particularly positive or negative opinions.

In this assignment, you will: (1) Apply sentiment analysis techniques to classify each review as positive, negative, or neutral; (2) Analyze the overall sentiment distribution across the dataset to evaluate user satisfaction; (3) Identify and examine the most positive and the most negative reviews to understand the reasons behind extreme reactions.

*!!! Now it's your turn:*

**Assignment 3** ˇV Apply sentiment analysis to the user reviews to assess their polarity, and identify the most positive and most negative entries for deeper insight into consumer reactions.

```{r}

```

## **Assignment 4**: *Extracting Smarphone Components in User-reviews*

In this assignment, we shift from general text analysis to a more focused **extraction task** aimed at identifying **key smartphone components** mentioned in user reviews.

The goal is to extract and analyze the main technological components discussed in the reviews by applying a **Named Entity Recognition (NER)** approach using a gazetteer specific to smartphone parts. This will help us understand which components users talk about most frequently and highlight the features that matter most in consumer feedback.

Follow these steps: (1) **Prepare a Gazetteer**: create a list of relevant smartphone components that you expect to find in the texts; (2) **Apply Gazetteer-based NER**: Use the gazetteer to identify and extract mentions of these components from the text of the user-reviews; (3) **Analyze the Results**: Count how often each component appears and Identify the most frequently mentioned components. Provide a summary table or visualization of components frequency.

*Now it's your turn:*

Assignment 4 ˇV Create a **gazetteer of smartphone components**, extract them from user reviews using Named Entity Recognition, and analyze which components occur most frequently.

```{r}

```

## **Assignment 5**: *Mining* *Insights from Textual Data*

After exploring smartphone components mentioned in user reviews, we now aim to gain deeper insights by addressing key research questions about **user preferences and needs** based on the analyses previously performed.

The goal of this assignment is to **select one of the following research questions** and use the techniques learnedˇXsuch as text preprocessing, topic modeling, sentiment analysis, and entity extractionˇXto provide a data-driven answer. You are only expected to **design a clear plan** outlining how you would approach the analysis, and only **try to implement it [if you have time]{.underline}**.

1.  Which smartphone features are most appreciated by users?

2.  What latent customer needs emerge from the review analysis?

3.  Are there specific functionalities users would like to see in future smartphone models?

Think about the tools and methods introduced during the course to: (1) **Analyze the user reviews** with respect to your chosen question; (2) **Extract relevant insights** supported by quantitative and qualitative evidence; (3) **Summarize your findings** with clear interpretations and visualizations that illustrate the key points.

*!!! Now it's your turn:*

**Assignment 5** ˇV Choose [one]{.underline} research question related to smartphone user reviews, and **design a clear plan** outlining how you would approach the analysis.

```{r}

```
