---
title: "Project Work"
format: html
editor: visual
---

## Project Work

This notebook presents the project work developed as part of the course **"Text Analysis for Innovation Management"**, held on **July 3/4/5** in Pisa. The course was carried out by **Vito Giordano**, **Filippo Chiarello**, **Ivan Zupic** and **Irene Spada**.

The objective of this notebook is to provide guidelines in the application of **Natural Language Processing (NLP)** techniques to explore innovation-related topics, as introduced and practiced during the course. It includes the different **assignments and exercises** covered throughout the **hands-on sessions**.

By applying the methods discussed in class to a real-world case study, this work aims to demonstrate how textual data can be used to extract insights relevant to **innovation management** and decision-making.

## Case Study: Chatbot Patents

In recent years, the use of **chatbots** has grown exponentially, driven by advances in artificial intelligence, natural language processing, and machine learning. Chatbots are increasingly used across a wide range of sectorsĄXfrom customer service management to business process automation, as well as in entertainment and healthcare. This growth has sparked growing interest in innovation within this field, which is also reflected in the rising number of related **patents**. Analyzing chatbot-related patents provides an opportunity to identify the most **active areas of innovation** and understand **emerging technological trends**.

Innovation management is a crucial aspect for companies operating in high-tech markets. Through strategic innovation management, organizations can maintain their competitive edge and respond effectively to market changes. In this context, analyzing technological trends becomes an essential tool to **anticipate the future directions of innovation**, especially in fast-evolving sectors such as chatbots.

The hands-on component of this project is organized into **five assignments**, each focusing on a specific aspect of text analysis applied to the topic of **chatbot patents**:

-   **Assignment 1**: *Text Pre-processing of* *Chatbot Patents*

-   **Assignment 2**: *Topic Modeling on Chatbot Patents*

-   **Assignment 3**: *Sentiment Analysis on Chatbot Patents*

-   **Assignment 4**: *Extracting Technologies in Chatbot Patents*

-   **Assignment 5**: *Mining* *Insights from Textual Data*

## **Assignment 1**: *Text Pre-processing of* *Chatbot Patents*

The dataset contains a total of 2,101 patents (title and abstract only) collected using the R package **PatentsView**. The data were extracted by searching for the following keywords within the title and abstract of the patents:

`"chatbot" OR "chat bot" OR "chatterbot" OR "chatter bot" OR "chatterbot" OR "virtual assistant" OR "voice assistant" OR "chat gpt" OR "chatgpt" OR "intelligent personal assistant" OR "conversational agent" OR "conversational bot" OR "conversational user interface".`

The resulting dataset is stored in a **CSV file** and includes:

-   The **title**, and **abstract** of each patent;

-   **Metadata**, such as the patent number, date, year, kind, and type.

The main goal of this assignment is to **explore and analyze the textual content** of chatbot patents, focusing specifically on Part-of-Speech (PoS) tagging and word count.

To conduct this analysis, you will apply a selection of Natural Language Processing (NLP) techniques introduced in the previous course modules. The workflow includes: (1) Text Preprocessing (Tokenization, PoS tagging, Stopword removal, Lemmatization or Stemming); and (2) Basic text analysis through word frequency counts and PoS distribution.

This assignment serves as a preliminary exploration of the dataset, aiming to provide initial insights into the structure and linguistic patterns of the job postings content. It is intended as a first step before moving on to more advanced analyses in subsequent assignments.

```{r}
library(tidyverse)

# Import Scientific Articles
chatbot_patents <- read_csv("data/Chatbot Patents.csv")

# Have a glimpse in the dataset
glimpse(chatbot_patents)

# Main Variables
chatbot_patents %>% 
  select(patent_number, patent_date, patent_year, patent_title, patent_abstrac, patent_kind, patent_type) %>% 
  head(5)
```

*!!! Now it's your turn:*

**Assignment 1** ĄV Perform a word count to identify the most frequent terms, then apply PoS tagging to extract the most common bigrams and nouns, gaining an initial overview of the chatbot patents.

```{r}

```

## **Assignment 2**: *Topic Modeling on Chatbot Patents*

Building on the insights gained from Assignment 1, the focus of this task shifts from basic text preprocessing to deeper analysis through **topic modeling**.

The goal of this assignment is to uncover the **main technological themes** emerging in the development of chatbot technologies. By applying topic modeling techniques to chatbot-related patents, you will identify which **innovations, methods, and application** areas are most frequently addressed by inventors, offering a data-driven view of how chatbot technologies are evolving.

Use the cleaned and preprocessed data from Assignment 1 to: **(1) Apply topic modeling** to extract key topics from the chatbot patent dataset; **(2) Identify and visualize** the most representative terms for each topic; and **(3) Interpret and label** each topic in terms of technological trends, application domains, or innovation directions.

*!!! Now it's your turn:*

**Assignment 2** ĄV Based on the processed patent data from Assignment 1, apply topic modeling to uncover the main themes in chatbot-related patents and interpret the results.

```{r}

```

## **Assignment 3**: *Sentiment Analysis on Chatbot Patents*

After exploring linguistic patterns and key topics in chatbot-related patents, this assignment focuses on analyzing the **emotional tone** of innovation narratives through **sentence-level sentiment analysis**.

The goal is to assess how positively or negatively chatbot technologies are framed in patent texts by measuring the **distribution of sentiment** across titles and abstracts. This allows us to understand the tone of technological discourse, detect areas of excitement or concern, and identify whether certain innovations are presented more optimistically or cautiously.

In this assignment, you will: **(1) Apply sentiment analysis techniques** to classify each sentence from the patent titles and abstracts as positive, negative, or neutral; **(2) Analyze the overall sentiment** distribution to evaluate how chatbot innovations are described; **(3) Identify and examine the most positively and negatively framed sentences** to understand the narratives driving the extremes in sentiment.

*!!! Now it's your turn:*

**Assignment 3** ĄV Apply sentiment analysis to the sentences from chatbot-related patents to assess their polarity, and identify the most positive and most negative sentences.

```{r}

```

## **Assignment 4**: *Extracting Technologies in Chatbot Patents*

In this assignment, we shift from general text analysis to a more focused extraction task aimed at identifying **key technologies mentioned in chatbot-related patents**.

The goal is to extract and analyze the main technical elements discussed in patent titles and abstracts by applying a **Named Entity Recognition (NER)** approach using a gazetteer specific to chatbot technologies. This will help us understand which components, methods, or systems are central to innovation in this domain, and highlight the technological trends shaping the development of chatbot applications.

To simplify the NER task, we will use a **gazetteer-based approach**, relying on a curated list of technology terms. Moreover, use the following definition of technology:

Technology

:   A technology is a technical mean or in general a technical system (A) created by human-kind (B) through the application of knowledge and science (C) in order to solve a practical problem or perform a function (D).

Follow these steps: **(1) Prepare a Gazetteer**: create a list of relevant AI technologies that you expect to find in the texts; **(2) Apply Gazetteer-based NER**: Use the gazetteer to identify and extract mentions of these technologies from the chatbot patents; **(3) Analyze the Results**: Count how often each item appears and identify the most frequently mentioned technologies. Provide a summary table or visualization of frequency data.

*!!! Now it's your turn:*

**Assignment 4** ĄV Create a gazetteer of AI technologies, extract them from chatbot patents using Named Entity Recognition, and analyze which ones occur most frequently.

```{r}

```

## **Assignment 5**: *Mining* *Insights from Textual Data*

After exploring the technologies and innovation patterns found in chatbot-related patents, we now aim to gain deeper insights by addressing **key research questions** about the evolution of this technological field, based on the analyses previously performed.

The goal of this assignment is to **select one** of the following research questions and use the techniques learnedĄXsuch as text preprocessing, topic modeling, sentiment analysis, and entity extractionĄXto provide a data-driven answer. You are only expected to **design a clear plan** outlining how you would approach the analysis, and only **try to implement it [if you have time]{.underline}**.

1.  Which technologies are most frequently mentioned in chatbot-related patents, and how are they evolving over time?

2.  What are the main application areas and use cases for chatbots as reflected in patent filings?

Think about the tools and methods introduced during the course to: (1) Analyze the patent dataset with respect to your chosen question; (2) Extract relevant insights supported by both quantitative and qualitative evidence; (3) Summarize your findings with clear interpretations and visualizations that highlight the main takeaways.

*!!! Now it's your turn:*

**Assignment 5** ĄV Choose [one]{.underline} research question related to smartphone user reviews, and **design a clear plan** outlining how you would approach the analysis.

```{r}

```
